---
title: "Untitled"
output: pdf_document
---


# Setup

## Packages and Data

```{r}
library(tidyverse)

rm(list = ls())

source("0 - Functions - Simulation.R")

load("Data/Simulation Data/Sim Data.Rdata")

# ggplot(df.clusters, aes(UCS_Rank, SCS_Rank, color = factor(strata))) + 
#   geom_point() + 
#   facet_wrap(~ K) + 
#   theme_minimal()


```


## Specify Parameters and inputs

```{r}

sim.data <- df.sim %>% 
  select(DSID, all_of(covariates))

B.index.covariates <- c(covariates[!(covariates %in% c("Urban", "Suburban", "ToRu"))], "urbanicity")

B.index.formula <- as.formula(paste("accepted ~", paste(B.index.covariates, collapse = " + ")))

B.index.data <- df.sim %>%
  select(DSID, all_of(B.index.covariates))

df.K.ranks <- df.clusters %>%
              nest(cluster.ranks = -c(K)) %>%
              rename(K.condition = K)

params <- list(
  iterations = 12,
  RR.condition = unique(df.PS$RR),
  K.condition = c("K_05"),
  SB.condition = unique(df.PS$scale_factor)
)


param_df <- df.PS %>%
  nest(PS.data = -c(scale_factor, RR)) %>%
  rename(SB.condition = scale_factor,
         RR.condition = RR) %>%
  right_join(cross_df(params)) %>%
  left_join(df.K.ranks) %>%
  mutate(B.index.formula = list(B.index.formula),
         B.index.data = list(B.index.data),
         sim.data = map(cluster.ranks, function(x) select(x, strata, DSID) %>% left_join(sim.data)))



```

# Function Test

```{r}

test <- param_df %>% filter(SB.condition == 1 ,
                            RR.condition %in% c("RR_20", "RR_50"))

PS.data <- test$PS.data[[1]]
cluster.ranks <- test$cluster.ranks[[1]]
sim.data <- test$sim.data[[1]]

```

## Generate Potential Responses

```{r}


df.responses <- Generate_Responses(PS.data, cluster.ranks)

responses.multitest <- test %>%
  mutate(df.responses = pmap(.l = list(PS.data = PS.data, 
                               cluster.ranks = cluster.ranks), 
                     .f = Generate_Responses)) %>%
  select(SB.condition, RR.condition, K.condition, df.responses)




```

## Select Samples

```{r}
df.sampled <- Create_Samples(df.responses)


sampled.multitest <- responses.multitest %>%
  mutate(df.sampled = map(df.responses, Create_Samples)) %>%
  select(-df.responses)



```


## Calculate Statistics

### Recruitment Statistics
```{r}

df.recruitment.stats <- Calc_Recruitment_Stats(df.sampled) 

rstats.multitest <- sampled.multitest %>%
  mutate(df.recruitment.stats = map(df.sampled, Calc_Recruitment_Stats)) %>%
  select(-df.sampled)


```
### Track Sampled Schools
```{r}

df.samp.counts <- df.sampled %>%
  filter(accepted) %>%
  select(sample_method, DSID) 


sampcounts.multitest <- sampled.multitest %>%
  mutate(df.samp.counts = map(df.sampled, function(x) filter(x, accepted) %>% select(sample_method, DSID))) %>%
  select(-df.sampled)



```

### Means and SDs
```{r}
df.samp.stats <- df.samp.counts %>%
  left_join(sim.data) %>%
  Calc_Sample_Statistics

sampstats.multitest <- sampcounts.multitest %>%
  mutate(df.samp.stats = map(df.samp.counts, function(x) left_join(x, sim.data) %>% Calc_Sample_Statistics)) %>%
  select(-df.samp.counts)


```

### B-Index
```{r}


df.B.indicies <- Bindex_Summary(df.sampled, B.index.data, B.index.formula)

bindex.multitest <- sampled.multitest %>%
  mutate(df.B.indicies = map(df.sampled, Bindex_Summary, B.index.data, B.index.formula)) %>%
  select(-df.sampled)


```


```{r}
test.results <- test %>%
  mutate(df.responses = pmap(.l = list(PS.data = PS.data, 
                               cluster.ranks = cluster.ranks), 
                     .f = Generate_Responses),
         df.sampled = map(df.responses, Create_Samples),
         df.recruitment.stats = map(df.sampled, Calc_Recruitment_Stats),
         df.samp.counts = map(df.sampled, function(x) filter(x, accepted) %>% select(sample_method, DSID)),
         df.samp.stats = map2(df.samp.counts, sim.data, function(x, y) left_join(x, y) %>% Calc_Sample_Statistics),
         df.B.indicies = pmap(.l = list(df.sampled, B.index.data, B.index.formula), .f = Bindex_Summary)) %>%
  select(SB.condition, RR.condition, K.condition, df.recruitment.stats:df.B.indicies)



test.results

test.results$df.recruitment.stats[[1]]
test.results$df.samp.counts[[1]]
test.results$df.samp.stats[[1]]
test.results$df.B.indicies[[1]]
```


# Iteration Test

## Setup

### Packages and Data

```{r}
library(tidyverse)

rm(list = ls())

source("0 - Functions - Simulation.R")

load("Data/Simulation Data/Sim Data.Rdata")

# ggplot(df.clusters, aes(UCS_Rank, SCS_Rank, color = factor(strata))) + 
#   geom_point() + 
#   facet_wrap(~ K) + 
#   theme_minimal()


```


### Specify Parameters and inputs

```{r}

sim.data <- df.sim %>% 
  select(DSID, all_of(covariates))

B.index.covariates <- c(covariates[!(covariates %in% c("Urban", "Suburban", "ToRu"))], "urbanicity")

B.index.formula <- as.formula(paste("accepted ~", paste(B.index.covariates, collapse = " + ")))

B.index.data <- df.sim %>%
  select(DSID, all_of(B.index.covariates))

df.K.ranks <- df.clusters %>%
              nest(cluster.ranks = -c(K)) %>%
              rename(K.condition = K)

params <- list(
  iterations = 2,
  RR.condition = unique(df.PS$RR),
  K.condition = c("K_05"),
  SB.condition = unique(df.PS$scale_factor)
)


param_df <- df.PS %>%
  nest(PS.data = -c(scale_factor, RR)) %>%
  rename(SB.condition = scale_factor,
         RR.condition = RR) %>%
  right_join(cross_df(params)) %>%
  left_join(df.K.ranks) %>%
  mutate(B.index.formula = list(B.index.formula),
         B.index.data = list(B.index.data),
         sim.data = map(cluster.ranks, function(x) select(x, strata, DSID) %>% left_join(sim.data)))


test <- param_df %>% filter(SB.condition == 1 ,
                            RR.condition %in% c("RR_20", "RR_50"))

```

## Iteration Function

```{r}


Run_Iteration(test$PS.data[[1]], 
              test$cluster.ranks[[1]], 
              test$sim.data[[1]], 
              test$B.index.data[[1]], 
              test$B.index.formula[[1]])

test %>%
  mutate(results = pmap(.l = list(PS.data, cluster.ranks, sim.data, B.index.data, B.index.formula),
                        .f = Run_Iteration))


# undebug(Run_Iteration)

```

## Iteration Timer

```{r}
Time_Iteration <- function(x = 1, PS.data, cluster.ranks, sim.data, B.index.data, B.index.formula) {
  system.time(Run_Iteration(x = x, PS.data, cluster.ranks, sim.data, B.index.data, B.index.formula))[3] / 60 
}
```


### Partial Conditions

```{r}


test.time <-
  Time_Iteration(x = 1,
                 test$PS.data[[1]], 
                 test$cluster.ranks[[1]], 
                 test$sim.data[[1]], 
                 test$B.index.data[[1]], 
                 test$B.index.formula[[1]])

system.time(multitest.time <- 
  test %>%
  mutate(minutes = pmap(.l = list(x = 1, PS.data, cluster.ranks, sim.data, B.index.data, B.index.formula),
                        .f = Time_Iteration)) %>%
  select(SB.condition, RR.condition, K.condition, minutes) %>% 
  unnest(minutes) %>%
  mutate(seconds = minutes * 60))

multitest.time
```

### All Conditions

```{r}

multitest.time <- 
  param_df %>%
  mutate(minutes = pmap(.l = list(x = 1, PS.data, cluster.ranks, sim.data, B.index.data, B.index.formula),
                        .f = Time_Iteration)) %>%
  select(SB.condition, RR.condition, K.condition, minutes) %>% 
  unnest(minutes) %>%
  mutate(seconds = minutes * 60)

multitest.time
```

# Parallel Test

```{r}
Sim_Driver  <- function(iterations, PS.data, cluster.ranks, sim.data, B.index.data, B.index.formula) {

  results <- map_dfr(
    .x = 1:iterations,
    .f = Run_Iteration,
    PS.data = PS.data, 
    cluster.ranks = cluster.ranks, 
    sim.data = sim.data,
    B.index.data = B.index.data, 
    B.index.formula = B.index.formula
  )
  
  r.stats <- 
    results$df.recruitment.stats %>%
    bind_rows() %>%
    group_by(sample_method, strata) %>%
    summarise_all(mean)
  
  samp.counts <-
    results$df.samp.counts %>%
    bind_rows() %>%
    group_by(sample_method, DSID) %>%
    count()
  
  B <- results$df.B.indicies %>%
    bind_rows()
  
  smd.stats <- results$df.samp.stats %>%
    bind_rows() %>%
    group_by(sample_method, strata, var) %>%
    summarise(sim.mean = mean(samp.mean),
              sim.sd = sd(samp.mean))
  
  tibble(r.stats = list(r.stats),
         samp.counts = list(samp.counts), 
         B = list(B),
         smd.stats = list(smd.stats))

}

test.results <- Sim_Driver(iterations = 2,
                           PS.data = test$PS.data[[1]],
                           cluster.ranks = test$cluster.ranks[[1]], 
                           sim.data = test$sim.data[[1]], 
                           B.index.data = test$B.index.data[[1]], 
                           B.index.formula = test$B.index.formula[[1]])




```

```{r}
library(future)
library(furrr)
plan(multisession)

system.time(multitest.results <- test %>%
    mutate(minutes = future_pmap(.l = list(x = 1, PS.data, cluster.ranks, sim.data, B.index.data, B.index.formula),
                             .f = Time_Iteration,
                             .options = furrr_options(seed = NULL))) %>%
  select(SB.condition, RR.condition, K.condition, minutes) %>% 
  unnest(minutes) %>%
  mutate(seconds = minutes * 60))

```

```{r}
test$iterations <- 2

sim.test.results <- 
  system.time(multitest.results <- test %>%
    mutate(results = future_pmap(.l = list(iterations, PS.data, cluster.ranks, sim.data, B.index.data, B.index.formula),
                             .f = Sim_Driver,
                             .options = furrr_options(seed = NULL))) %>%
      select(SB.condition, RR.condition, K.condition, results))

(sim.test.results[[3]] / 2 * 1000) / 60 / 60

```

```{r}

param_df$iterations <- 2

sim.test.results <- 
  system.time(multitest.results <- param_df %>%
    mutate(results = future_pmap(.l = list(iterations, PS.data, cluster.ranks, sim.data, B.index.data, B.index.formula),
                             .f = Sim_Driver,
                             .options = furrr_options(seed = NULL))) %>%
      select(SB.condition, RR.condition, K.condition, results))

(sim.test.results[[3]] / 2 * 1000) / 60 / 60

multitest.results %>%
  unnest(results)
```


