---
title             : "GenSamp: RESULTS"
shorttitle        : "RESULTS"

author:
  - name          : "Gleb Furman"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Texas at Austin, SZB 504, 1912 Speedway, Austin, Texas 78712"
    email         : "gleb.furman@utexas.edu"
  - name          : "James E. Pustejovsky"
    affiliation   : "2"
  - name          : "Elizabeth Tipton"
    affiliation   : "3"
    

affiliation:
  - id            : "1"
    institution   : "University of Texas at Austin"
  - id            : "2"
    institution   : "University of Wisconsin-Madison"
  - id            : "3"
    institution   : "Northwestern University"


abstract: |
    <!-- In order for educational research to be informative to policy makers, studies must be designed to make robust estimates of causal effects at the population level. Large scale multi-site randomized trials (MRT) often rely on vague convenience sampling methodology when recruiting districts and schools, resulting in relatively homogeneous samples that may differ greatly from the intended population of interest. Retrospective methods that quantify and statistically adjust for those differences are promising but have limited effect when the sample differs greatly from the population. Designing sampling methods that focus on generalizability may be a more effective but costly solution, but limited methodological research has been performed to examine their effectiveness in the educational context. This paper examines one promising method, stratified balanced sampling (SBS), in the context of recruiting a representative sample of schools for a large scale MRT. Using simulations based on real data, we compare SBS to stratified and unstratified versions of convenience sampling and probability sampling. Several models for generating school participation and emulating convenience sampling are proposed. Results indicate that SBS and stratified random sampling (SRS) result in highly generalizable samples. These methods are extremely costly to implement, however, especially when the population average willingness to participate is low. Stratified convenience sampling (SCS) is a potential compromise. -->

# keywords          : "keywords"
# wordcount         : "X"

bibliography      : ["references.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---


```{r setup, include = FALSE}
rm(list = ls())

output.screen <- T

knitr::opts_chunk$set(echo = F,
                      cache = F,
                      message = output.screen,
                      warning = output.screen,
                      results = "asis",
                      knitr.table.format = "latex")
options(knitr.kable.NA = "-")

```

# Setup

## Packages and Data

```{r load.packages, include = FALSE}
library(papaja)
library(tidyverse)
library(kableExtra)
library(ggalt)
library(ggrepel)
library(ggforce)
library(concaveman)
library(cowplot)

rm(list = ls())

load(file = "Data/Paper/Summarised Sim Results.RData")

source("0 - Functions - Analysis and Paper.R")

```


## Organize Objects

```{r}

grey.plots <- F


lab.vars <- list(k = "Number of strata (k)", 
                 vratio = expression(paste("Variance Ratio (",
                                           sigma[B]^2/sigma[T]^2,
                                           ")")), 
                 ch = expression("CH index (" ~ 10^3 ~ ")"),
                 n_j = "Target sample per cluster",
                 pop.prop = "Proportion of population (%)",
                 smd = "Standardized mean difference",
                 vars = "Participation covariates",
                 intercept = "Participation model intercept",
                 rr = "Population participation \n(generated %)",
                 B = "Average B-index",
                 coef = "Participation model coeficients \n(absolute log-odds)",
                 contacts = expression("Schools contacted (" ~ 10^3 ~ ")"),
                 rr.o = "Population participation  \n(observed %)")
```

# Data Summary


## Covaraite Statistics
```{r}
tabs$pop.desc <- 
  df.list$pop.stats %>%
  ungroup() %>%
  arrange(Category, Sub) %>%
  select(Category, Variables, pop.mean, pop.sd, log_odds) %>%
  mutate(M = round(pop.mean, 2),
         SD = round(pop.sd, 2),
         .keep = "unused")

```

## Continuous variable distributions
```{r}
tabs$dist <- 
  df.list$pop.data %>%
  select(all_of(vars$cluster$mixed)) %>%
  select_if(is.numeric) %>%
  mutate(n = log(n),
         ST.ratio = log(ST.ratio),
         dSCH = log(dSCH)) %>%
  gather(key = vnames, value = val) %>%
  left_join(df.list$coefs,
            by = "vnames")

figs$dist <- 
  tabs$dist %>%
  ggplot(aes(x = val)) +
  geom_histogram() +
  scale_y_continuous(labels = function(x) x / 100) +
  facet_wrap(~ Variables, scales = "free", ncol = 3) +
  labs(x = "", y = "Schools (10^2)")

figs$dist

```


# Methods Summary

## Cluster Analysis

### Selecting k
```{r}
hlines <- 
  df.list$k.stats %>%
  filter(K %in% c(2, 4)) %>%
  select(ch) %>%
  unlist

figs$ch <- 
  df.list$k.stats %>%
  ggplot(aes(x = K, y = ch)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = hlines, linetype = "dotted") +
  scale_x_continuous(breaks = c(2:10)) +
  scale_y_continuous(breaks = seq(5000, 8500, 1000),
                     limits = c(5000, 8500),
                     labels = function(x) x / 1000) +
  labs(y = lab.vars$ch,
       x = lab.vars$k)

figs$ch
```

```{r}



figs$vratio <- 
  df.list$k.stats %>%
  bind_rows(summarise(., K = 1, bt.ss = 0)) %>%
  mutate(min80 = sum(bt.ss <= .8) + min(K) - .5) %>%
  ggplot(aes(x = K, y = bt.ss)) +
  geom_point() +
  geom_vline(aes(xintercept = min80), linetype = "dotted") +
  geom_hline(yintercept = .8, linetype = "dotted") +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(breaks = seq(0, 1, .25), limits = c(0, 1)) +
  geom_line() +
  labs(y = lab.vars$vratio,
       x = lab.vars$k)

figs$vratio
```

```{r}




tabs$allocation <- 
  df.list$cluster.pa %>%
  mutate(k = parse_number(K),
         strata = paste("Strata", strata))

figs$allocation <- 
  tabs$allocation %>%
  ggplot(aes(x = k, y = pa)) +
  geom_point(shape = "-", size = 15, alpha = .5) +
  labs(y = lab.vars$n_j,
       x = lab.vars$k) +
  scale_x_continuous(breaks = seq(2, 10, 1)) +
  scale_y_continuous(breaks = seq(0, 50, 5),
                     limits = c(0, 35),
                     sec.axis = sec_axis(~./.6, 
                                         name = lab.vars$pop.prop,
                                         breaks = seq(0, 100, 10))) +
  stat_function(fun = function(x) 60/x, geom = "line", linetype = "dashed", size = .5) +
  geom_hline(yintercept = 5, linetype = "dotted", size = .5)

figs$allocation

```

```{r}

figs$joint.cluster.stats <- 
  plot_grid(plotlist = list(figs$ch,
                            figs$vratio), 
          labels=c("(a)", "(b)"), 
          ncol = 2, 
          nrow = 1, 
          rel_heights = c(1, 1),
          label_x = .5,
          align = "v",
          label_fontface = "plain" ,
          label_size = 10) %>%
  plot_grid(figs$allocation, 
            ncol = 1,
            nrow = 2, 
            labels = c("", "(c)"),
            label_x = .5,
            # align = "h",
            label_fontface = "plain" ,
            label_size = 10)

figs$joint.cluster.stats 
```

```{r, results = "asis"}

tabs$cluster.SMD <- 
  df.list$pop.data %>% 
  select(strata, all_of(vars$cluster$dummy)) %>%
  gather(key = vnames, value = val, -strata) %>%
  group_by(strata, vnames) %>%
  summarise(strat.mean = mean(val), .groups = "drop_last") %>%
  left_join(df.list$pop.stats, by = "vnames") %>%
  transmute(vnames = vnames,
            SMD = round((pop.mean - strat.mean) / pop.sd, 2)) %>%
  left_join(df.list$coefs, by = "vnames")

```


```{r}


figs$cluster.SMD <- 
  tabs$cluster.SMD %>%
  mutate(sig = ifelse(abs(SMD) > .25, "Sig", "Not Sig"),
         neg = SMD > 0,
         k = parse_number(as.character(strata))) %>%
  ggplot(aes(x = SMD, y = Variables, color = strata, alpha = sig, shape = strata)) +
  geom_point() +
  geom_vline(xintercept = c(-.25, .25), linetype = "dotted") +
  scale_alpha_discrete(guide = "none")  +
  scale_x_continuous(limits = c(-2,2)) +
  # theme(axis.text.y = element_blank(),
  #       axis.ticks.y = element_blank()) +
  labs(x = lab.vars$smd ,
       y = lab.vars$vars)

figs$cluster.SMD
```





## Variation explained by the strata
```{r}
tabs$strata.data <- 
  tabs$allocation %>%
  filter(k == 5) %>%
  left_join(df.list$k.stats %>% filter(K == 5) %>%
              rename(k = K), by = "k") %>%
  mutate(bet.perc = round(bt.ss * 100, 2),
         ch = round(ch, 2))




```

## Participation Generating Model

```{r}

 
df.list$coefs %>% 
  select(Sub, Category, Type, Variables, log_odds) %>%
  unique() %>%
  kable() %>%
  kable_styling()


```


```{r}


tabs$intercepts <- 
  intercepts %>%
  mutate(RR = parse_number(RR),
         scale_factor = factor(scale_factor),
         Intercept = Int)

tabs$intercepts %>%
  ggplot(aes(x = RR, y = Intercept, group = scale_factor, color = scale_factor)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(0, 100, 10)) +
  labs(y = lab.vars$intercept,
       x = lab.vars$rr)
```




```{r, fig.cap = "Distributions of Participation Logits"}

df.list$PS %>%
  mutate(RR = as_factor(RR)) %>%
  ggplot(aes(x = logit, y = RR, fill = RR)) +
  ggridges::geom_density_ridges(stat = "binline", 
                                bins = 100, 
                                alpha = .3, 
                                scale = 2) +
  theme_apa() +
  theme(legend.position = "none")+
  scale_x_continuous(breaks = seq(-8,10, 2))  +
  facet_wrap(~scale_factor)
```

# Results

## Generalizability

### B Index

```{r}

get_quant <- function(x) {
  tibble(
    Q = quantile(x), 
    Percent = paste("q_", 
                    parse_number(names(Q)), 
                    sep = "")) %>% 
    spread(Percent, Q)
}

set_compare <- function(data, variable = "Sampler", 
                        value = "S_BS_SR_X1", 
                        observations = c("M", "q_25", "q_50", "q_75")){
  data %>%
  filter(.[[variable]] == value) %>%
  nest(Comparison = c(variable, observations)) %>%
  select(SB:K,
         Comparison) %>%
  mutate(Comparison = map(Comparison, rename_with, function(x) paste(x, "_C", sep = ""))) %>%
  unnest(Comparison)
}

tabs$B <- 
  results$B %>%
  chop(Bs) %>%
  mutate(M = map_dbl(Bs, mean),
         SD = map_dbl(Bs, sd),
         LL = M - SD,
         UL = M + SD,
         Quantile = map(Bs, get_quant)) %>%
  select(-Bs) %>%
  unpack(Sample_Catagories) %>%
  unnest(Quantile) 


  
```

```{r, fig.cap = "Averge B-Index"}
figs$B <- tabs$B %>%
  filter(!(Rank_Factor %in% c("X9"))) %>%
  ggplot(aes(x = RR, y = M)) %>%
  common_line() %>%
  common_point() +
  geom_hline(yintercept = c(.8, .95), linetype = "dotted", alpha = .5) +
  theme_minimal() +
  facet_wrap(~SB) +
  labs(y = lab.vars$B,
       x = lab.vars$rr)

figs$B
```


```{r}
tabs$B %>%
  left_join(set_compare(.), by = c("SB", "RR", "K")) %>%
  filter(SB == 1,
         Sampler != Sampler_C,
         !(Rank_Factor %in% c("X9"))) %>%
  ggplot(aes(x = RR, y = M, group = Sampler, color = Cluster, fill = Cluster)) +
  geom_line() +
  geom_ribbon(aes(ymin = q_25, ymax = q_75), alpha = .1, linetype = "dotted") +
  geom_line(aes(y = M_C), color = "black", alpha = .2) +
  geom_ribbon(aes(ymin = q_25_C, ymax = q_75_C), alpha = .1, linetype = "dotted", color = "black") +
  geom_hline(yintercept = c(.8, .95), linetype = "dotted", alpha = .5) +
  facet_wrap(~ Sample + Rank_Factor, nrow = 1) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(y = lab.vars$B,
       x = lab.vars$rr)
  

```




### Standardized mean differences

```{r, fig.cap = "Average SMDs", fig.height = 7}

tabs$smd.results <- 
  results$smd.stats %>%
  select(any_of(vars$methods), any_of(vars$conditions),
         strata, log_odds, mSMD, abs.mSMD) %>%
  unpack(Sample_Catagories) %>%
  filter(!(Rank_Factor %in% c("X9")),
         strata == 0) %>%
  select(-strata, -log_odds) %>%
  mutate(Value = abs.mSMD,
         alpha.group = as.numeric(Value > .25))

tabs$smd.by.scale <- tabs$smd.results %>%
  nest(smd.plot.data = -c(Variables, vnames)) %>%
  mutate(plots = pmap(.l = list(data = smd.plot.data, g.title = Variables), .f = plot_smd, l.pos = "bottom"),
         plots = setNames(plots, vnames))

figs$smd.by.scale <- tabs$smd.by.scale$plots %>% as.list




```



```{r}

figs$smd.by.scale$Suburban
figs$smd.by.scale$T1
figs$smd.by.scale$Urban


```

```{r}


figs$smd.by.scale$pTotfrl
figs$smd.by.scale$ethHisp
figs$smd.by.scale$ethWhite
figs$smd.by.scale$pELL
figs$smd.by.scale$dSCH
```


```{r}

figs$smd.by.scale$n
figs$smd.by.scale$ethBlack
figs$smd.by.scale$ToRu

```

```{r}

figs$smd.by.scale$pFem
figs$smd.by.scale$ST.ratio

```




```{r}

vars$smd.groups <- 
  tibble(group1 = list(c("Suburban", "T1", "Urban")),
         group2 = list(c("pTotfrl", "ethHisp", "ethWhite", "pELL", "dSCH")),
         group3 = list(c("n", "ethBlack", "ToRu")),
         group4 = list(c("pFem", "ST.ratio"))) %>%
  pivot_longer(cols = everything(), names_to = "Group", values_to = "vnames") %>%
  unnest(vnames)



vars$smd.examples <-c("Suburban", "dSCH", "n", "ST.ratio")
```





```{r}
tabs$smd.examples <- tabs$smd.results %>%
  left_join(vars$smd.groups, by = "vnames") %>%
  filter(vnames %in% vars$smd.examples) 

smd.examples.sep <-
  tabs$smd.examples %>%
  nest(data = -c(Variables, Group)) %>%
  mutate(plots = map(data, plot_smd, l.pos = "none"))

# smd.examples.sep$plots[[1]]


figs$smd.examples <- tabs$smd.examples %>%
  plot_smd() +
  facet_grid(Variables ~ SB) 


figs$smd.examples
```



```{r}

df.list$smd.by.strata.results <- 
  results$smd.stats %>%
  # select(SB:vnames, Variables, log_odds, Sample_Catagories, mSMD, abs.mSMD) %>%
  unpack(Sample_Catagories) %>%
  filter(!(Rank_Factor %in% c("X9")),
         Cluster == "Stratified",
         strata != 0) %>%
  select(-K, -log_odds) %>%
  mutate(Value = abs.mSMD,
         alpha.group = as.numeric(Value > .25))

df.list$smd.by.strata.results %>%
  filter(SB == 1,
         vnames %in% c("ethBlack", "n")) %>%
  plot_smd() +
  facet_grid(Variables ~ strata)

```
### Examples for presentations

```{r, fig.cap = "Average SMDs", fig.height = 7}


# vars$plot.smd.examples <- bind_rows(
#   tibble(group = "Example", 
#          desc = "each", 
#          vnames = c("pELL", "n", "pFem")),
#   tibble(group = "ELL", 
#          desc = "ell", 
#          vnames = c("pELL")),
#   tibble(group = "N", 
#          desc = "n", 
#          vnames = c("n")),
#   tibble(group = "Fem", 
#          desc = "fem", 
#          vnames = c("pFem"))
#   )
# 
# SMDFigs <-  results$s.stats %>%
#   left_join(df.list$pop.stats) %>%
#   mutate(smd = (samp.mean - pop.mean) / pop.sd,
#          RR = parse_number(RR)) %>%
#   group_by(Sampler, RR, var, Sampling, Clustering, K) %>%
#   summarise(mSMD = mean(abs(smd)),
#             mSMD.sd = sd(abs(smd))) %>%
#   rename(vnames = var) %>%
#   left_join(df.list$coefs) %>%
#   ungroup() %>%
#   left_join(vars$plot.smd.examples ) %>%
#   group_by(group, desc) %>%
#   nest %>%
#   mutate(plots = map(data, plot_smd),
#          plots = set_names(plots, paste("smd.", desc, sep = "")),
#          data = set_names(data, paste("smd.", desc, sep = ""))) %>%
#   ungroup() %>%
#   select(data, plots)
# 
# tabs <- c(tabs, SMDFigs$data)
# figs <- c(figs, SMDFigs$plots)
# 
# figs$smd.ell
```

### V-ratio and Log odds


```{r}

# tabs$v.coef <- df.list$pop.data %>%
#   select(DSID, all_of(covariates), strata) %>%
#   gather(key = vnames, value = value, covariates) %>%
#   group_by(vnames) %>%
#   mutate(T.SS = var(value) * (n() - 1)) %>%
#   group_by(vnames, T.SS, strata) %>%
#   summarise(W.SS = var(value) * (n() - 1)) %>%
#   summarise(W.SS = sum(W.SS)) %>%
#   mutate(B.SS = T.SS - W.SS,
#          vratio = B.SS / T.SS) %>%
#   left_join(df.list$coefs) %>%
#   left_join(vars$plot.groups) %>%
#   mutate(SB_11 = log_odds,
#          SB_12 = log_odds * .5,
#          SB_14 = log_odds * .25) %>%
#   gather(key = factor, value = coef, SB_11:SB_14) %>%
#   mutate(abs_log_odds = abs(coef),
#          odds = exp(coef),
#          prob = odds / (1 + odds))%>%
#   arrange(group)
# 
# figs$v.coef <- tabs$v.coef  %>%
#   ggplot(aes(x = vratio, y = abs_log_odds, color = group, fill = group)) +
#   geom_mark_rect(expand = 0.04, color = "black") +
#   geom_point(size = 2) +
#   # geom_encircle(alpha = .2, s_shape = 0, expand = .1, spread = .05, color = "black") +
#   # geom_text_repel(aes(label = Variables), point.padding = .2, color = "black") +
#   labs(y = lab.vars$coef,
#        x = lab.vars$vratio) +
#   scale_x_continuous(breaks = seq(0, 1, .1)) +
#   scale_y_continuous(breaks = seq(0, .6, .1),
#   limits = c(0, .6)) +
#   theme(legend.position =)
# 
# figs$v.coef
```


## Feasibility

### Sampling Difficulty

```{r}
tabs$samples <- 
  results$r.stats %>% 
  unpack(Sample_Catagories) %>%
  filter(strata == 0,
         !(Rank_Factor %in% c("X9"))) %>%
  mutate(response_rate = accepted/contacted)


```


```{r, fig.cap = "Recruitment Counts"}
# figs$samples <- tabs$samples  %>%
#   ggplot(aes(x = RR, y = value, group = Sampler, color = Sampling, linetype = Clustering)) +
#   geom_point() +
#   geom_line() +
#   facet_wrap( ~ measure, scales = "free_y") +
#   expand_limits(y = 0) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
# 
# figs$samples
# 
# tabs$samples.contacts.labels <- 
#   tabs$samples %>%
#   filter(RR == 10) %>%
#   arrange(contacted) %>%
#   ungroup() %>%
#   mutate(RR = 5)
```





```{r, fig.cap = "Schools Contacted"}

tabs$samples.contacts.labels <-  
  tabs$samples %>%
  filter(RR == 10,
         SB == 1) %>%
  arrange(contacted) %>%
  ungroup() %>%
  mutate(RR = 5,
         contacted = c(200, 1000, 2200, 2900, 3700))

figs$samples.contacts <- 
  tabs$samples %>%
  filter(SB == 1) %>%
  ggplot(aes(x = RR, y = contacted)) %>%
  common_line() %>%
  common_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = seq(0, 100, 10)) +
  scale_y_continuous(breaks = seq(0, 4500, 500),
                     limits = c(0, 4500),
                     labels = function(x) x / 1000) +
  labs(x = lab.vars$rr,
       y = lab.vars$contacts) +
  geom_label(data = tabs$samples.contacts.labels,
             aes(label = Method_Label, color = Sample))

figs$samples.contacts


```

```{r, fig.cap = "Sampling response rates"}



tabs$samples.response.rates.labels <-  
  tabs$samples %>%
  filter(RR == 10,
         SB == 1) %>%
  arrange(response_rate) %>%
  ungroup() %>%
  mutate(value = c(10, 25, 37, 50, 70),
         RR = 45)


figs$samples.response.rates <- 
  tabs$samples %>%
  filter(SB == 1) %>%
  mutate(value = response_rate * 100) %>%
  ggplot(aes(x = RR, y = value)) %>%
  common_line() %>%
  common_point() +
  expand_limits(y = 0) +
  theme(legend.position = "none") +
  labs(x = lab.vars$rr,
       y = lab.vars$rr.o) +
  scale_x_continuous(breaks = seq(0, 100, 10)) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  geom_label(data = tabs$samples.response.rates.labels,
             aes(label = Method_Label,
                 color = Sample))

figs$samples.response.rates


```




```{r}

legend <- get_legend(
  figs$samples.contacts +
  theme_apa(box = F) + 
  theme(legend.box.margin = margin(0, 0, 0, 12))
)

figs$joint.feasibility <- plot_grid(plotlist = list(figs$samples.contacts + theme(legend.position = "none"),
                                                    figs$samples.response.rates + theme(legend.position = "none")),
                                    labels=c("(a)", "(b)"),
                                    ncol = 1, 
                                    nrow = 2, 
                                    label_x = .5,
                                    align = "v",
                                    label_fontface = "plain",
                                    label_size = 10)
  # plot_grid(legend, rel_widths = c(3, .4))

figs$joint.feasibility 

```

### Relative Performance

```{r}
tabs$relative.counts <- 
  tabs$samples  %>%
  ungroup() %>%
  select(SB:strata, Method_Label, response_rate, -Sampler) %>%
  spread(key = Method_Label, value = response_rate) %>%
  right_join(tabs$samples, by = c("SB", "RR", "K", "strata")) %>%
  gather(key = Comp.Method, value = Comp.Value,
         SBS, SRS, SCS, URS, UCS) %>%
  mutate(Comp.Rate = response_rate / Comp.Value) 

figs$relative.counts <- 
  tabs$relative.counts %>%
  filter(SB == 1) %>%
  ggplot(aes(x = RR, y = Comp.Rate)) %>%
  common_line() %>%
  common_point() +
  facet_wrap(~ Comp.Method, scales = "free") 

figs$relative.counts

```

### Gini Plot

```{r}
lab.vars$gini <- "Gini Coeficient"

tabs$gini.coef <-
  results$gini %>%
  select(-count.data) %>%
  unpack(Sample_Catagories) %>%
  filter(!(Rank_Factor %in% c("X9"))) %>%
  group_by(RR, SB) %>%
  arrange(Gini_Coef) %>%
  mutate(y.pos = 1 - (1:n()/n()) / 1.5,
         x.pos = 1 - y.pos)

# tabs$gini.coef.labels <-
#   tabs$gini.coef %>%
#   filter(RR == 10) %>%
#   mutate(RR = 5) %>%
#   arrange(Gini_Coef) %>%
#   mutate(Gini_Coef = c(.5, .65, .75, .85, .95))
```


```{r}


figs$gini <- tabs$gini.coef %>%
  ggplot(aes(x = RR, y = Gini_Coef)) %>%
  common_line() %>%
  common_point() +
  scale_x_continuous(breaks = seq(0, 100, 20)) +
  scale_y_continuous(breaks = seq(0, 1, .1),
                     limits = c(0, 1), expand = c(0,0)) +
  labs(x = lab.vars$rr,
      y = lab.vars$gini) +
  facet_grid(~ SB)
  # geom_label(data = tabs$gini.coef.labels,
  #            aes(label = Method_Label,
  #                color = Sample))

figs$gini


```

```{r}

tabs$gini.count <-
  results$gini %>%
  select(SB:Sample_Catagories) %>%
  unpack(Sample_Catagories) %>%
  unnest(count.data)

figs$gini.coef <- 
  tabs$gini.count %>%
  filter(SB == 1,
         !(Rank_Factor %in% c("X9"))) %>%
  ggplot(aes(x = c.prop.pop, y = c.prop.sel))  %>%
  common_line() %>%
  common_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  geom_label(aes(label = round(Gini_Coef, 2), x = x.pos, y = y.pos), 
             data = tabs$gini.coef, size = 3) +
  facet_wrap(~RR) +  
  # (if (grey.plots) {scale_colour_grey(start = 0, end = .7)}) +
  labs(x = "Cumulative proportion of population",
         y = "Cumulative percentage of times sampled") +
  theme_apa()

ggsave(filename = "Figs/Presentation/GINI.jpg", dpi = 100, width = 10, height = 6)
```



```{r}
save(figs,
     lab.vars,
     results,
     tabs,
     vars,
     file = "Data/Paper/Paper Data 02-08-2021.rdata")

ls() %>%
  tibble(Vars = .) %>%
  mutate(MB = map(Vars, function(x) object.size(eval(parse(text = x))) %>% format(units = "MB", digits = 3))) %>%
  unnest(MB) %>%
  arrange(desc(parse_number(MB)))
```

