---
title             : "Analysis Write-up"
shorttitle        : "Analysis Write-up"

author:
  - name          : "Gleb Furman"
    affiliation   : "1"
  #   corresponding : yes    # Define only one corresponding author
  #   address       : "Postal address"
  #   email         : "my@email.com"
  # - name          : "Ernst-August Doelle"
  #   affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Who Kneads a PH.D. Bakery"
#   - id            : "2"
#     institution   : "Konstanz Business School"

# author_note: |
#   Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

#   Enter author note here.

# abstract: |
#   Enter abstract here. Each new line herein must be indented, like this line.
  
# keywords          : "keywords"
# wordcount         : "X"

# bibliography      : ["r-references.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache = F)
```

```{r load.packages, include = FALSE}
library(papaja)
library(tidyverse)

rm(list = ls())
```

```{r analysis_preferences}
# Seed for random number generation
# set.seed(42)
```

```{r data}
load("data/base data.rdata")

vars <- c("LSTATE", "LEANM", "SCHNAM", "DID", "SID", "DSID", "n", "pTotfrl",
          "Urban", "Suburban", "ToRu",
          "pELL", "pED", "pELA", "pMath", "pMin", "MEDINC", "MEDINC_SD")

# Covarites 
subs_f_vars <- c("n", "pTotfrl", "urbanicity",
                "pELL", "pED", "pELA", "pMath", "pMin", "MEDINC")

subs_ov_vars <- c("n", "pTotfrl", "urbanicity",
                "pELL", "pED", "pELA", "pMath", "pMin")

# School Level Data
df.sch <- df[,vars]


# District Level Data
df.dist <- select(df.sch, -SCHNAM, -LSTATE, -LEANM, -SID, -MEDINC_SD) %>%
  group_by(DID) %>%
  mutate(m = n()) %>%
  summarise_if(is.numeric, mean) %>%
  ungroup()


df.dist.w <- list(.vars = lst("n", vars[8:17], "m"),
     .funs = lst(mean, funs(weighted.mean(., w = n)), mean)) %>%
  pmap(~ df.sch %>%
         mutate(m = n()) %>%
         group_by(DID) %>%
         summarise_at(.x, .y)) %>%
  reduce(inner_join, by = "DID")


df.dist$dif <- df.dist.w$dif <- apply((df.dist[,-1] - df.dist.w[,-1])/apply(df.dist[,-1], 2, function(x) max(x) - min(x)), 1, function(x) mean(abs(x)))



```

```{r descriptives}
tbl_desc <- df.dist %>%
  gather(key = Variable, value = Value, n:MEDINC, m) %>%
  group_by(Variable) %>%
  summarise(M = mean(Value), SD = sd(Value))

tbl_desc <- df.dist.w %>%
  gather(key = Variable, value = Value, n:MEDINC, m) %>%
  group_by(Variable) %>%
  summarise(M = mean(Value), SD = sd(Value)) %>%
  right_join(tbl_desc, by = "Variable")

tbl_desc <- df.sch %>%
  gather(key = Variable, value = Value, n:MEDINC) %>%
  group_by(Variable) %>%
  summarise(M = mean(Value), SD = sd(Value))  %>%
  right_join(tbl_desc, by = "Variable")



tbl_desc$Variable <- c("Number of Schools", "Median Income", "School Size", "Economically Disadvantaged", "ELA Proficiency", "English Language Learners", "Math Proficiency", "Minority Status", "Total/Free/Reduced Lunch", "Suburban", "Town or Rural", "Urban")

tbl_desc <- tbl_desc[c(1,3,2,5,7,4,6,8,9,12,10,11),]


```

```{r tbl_desc, results = "asis"}

apa_table(tbl_desc,
          caption = "Descriptives of variables",
          note = "District variables are derived as aggregate means of school variables",
          stub_indents = list(`Average Proportions` = c(4:9), Indicators = c(10:12)),
          col_spanners = list(`School` = c(2,3), `District Weighted` = c(4,5), `District Unweighted` = c(6,7)),
          align = "lcccccc",
          col.names = c("Variables","Mean", "SD", "Mean", "SD", "Mean", "SD"))

```


# Cluster Analysis


## Population Frame
The population frame is composed of data from three sources: (1) the Common Core of Data (CCD), (2) publically available accountability data, and (3) the U.S. Census. The CCD is a comprehensive database housing anually collected national statistics of all public schools and districts. Accountability data was used to calculate the proportion of students within each school performing at or above proficiency in Math and ELA. Finally, local median income was obtained from the U.S. Census and was  matched to each school by zipcode. School level data was aggregated to get district level variables. These are reported in Table \@ref(tab:tbl_desc)

```{r distributions}
to_z <- function(x) (x - mean(x))/sd(x)

forDists <- df[, c("DSID",subs_f_vars[-2])] %>%
  mutate(n_z = to_z(n),
         MEDINC_z = to_z(MEDINC),
         n_log = log(n),
         MEDINC_log = log(MEDINC))

forDists %>%
  gather(key = var, value = value, n:MEDINC_log, -urbanicity) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~ var, scales = "free")

forDists %>%
  gather(key = var, value = value, n:MEDINC_log, -urbanicity) %>%
  ggplot(aes(x = value, color = urbanicity)) +
  geom_density() +
  facet_wrap(~ var, scales = c("free"))

forDists %>%
  ggplot(aes(x = pELA, y = pMath, color = n_log)) +
  geom_point(alpha = .5) +
  facet_wrap(~urbanicity)

forDists %>%
  ggplot(aes(x = pELL, y = pMin, color = n_log)) +
  geom_point(alpha = .5) +
  facet_wrap(~urbanicity)

forDists %>%
  ggplot(aes(x = MEDINC_log, y = pED, color = n_log)) +
  geom_point(alpha = .5) +
  facet_wrap(~urbanicity)
```

## SUBS

Stratification using balanced sampling (SUBS) was performed prior to simulation because the group of schools in each strata would be static across conditions except where the balancing model is manipulated. The set of covariates in both the full model (SUBS-F) and the omitted variable model (SUBS-OV) include binary indicator variables

```{r SUBS_Funcs}
library(parallel)
library(fpc)
library(cluster)
library(tidyverse)


run_clusters <- function(K, distance) {

  # # Calculate the number of cores
  # no_cores <- min(K, detectCores() - 7)
  # 
  # # Initiate cluster
  # cl <- makeCluster(no_cores)
  # 
  # clusters <- parLapply(cl, 1:K, function(x, y) kmeans(x = y, centers = x), distance)
  # 
  # stopCluster(cl)
  # 
  # # clusters <- bind_cols(clusters)
  # #
  # # names(classign) <- paste("k", 1:K, sep = "")
  
  clusters <- list()
  
  for(k in 1:K) {
    clusters <- append(clusters, list(kmeans(x = distance, centers = k)))
  }

  return(clusters)
}



get_CH <- function(K, distance, clusters) {
  ch <- list()
  
  for(k in 2:K) {
    ch1 <- (clusters[[k]]$betweenss / (k - 1)) / (clusters[[k]]$tot.withinss / (length(clusters[[k]]$cluster- k)))
    ch2 <- cluster.stats(distance, clusters[[k]]$cluster,
                                silhouette = F, G2 = FALSE, G3 = FALSE,
                                wgap=F, sepindex=F, sepprob=0.1,
                                sepwithnoise=F)
    ch_d <- ch2$ch
    
    ch = append(ch , list(k = k, 
                     ch = ch1,
                     ch_d = ch_d))
  }
  
  chPlot <- unlist(ch) %>%
  matrix(3, (K-1)) %>%
  t() %>%
  as.data.frame %>%
  mutate(dif = abs(V2 - V3))
  
  names(chPlot) <- c("k", "ch1", "ch2", "dif")

  
  return(chPlot)
}

```

### Number of Clusters
Selecting the number of clusters, $k$, is one of the most difficult problems in cluster analysis (Steinley, 2006). To date, the most extensive investigation of methods for determining $k$ was conducted by Milligan and Cooper (1985) who analyzed 30 methods. However, aside from the limited generalizability of this study, many methods are also innapropriate in the context of non-hierarchical and thus do not support k-means clustering. Hennig and Liao (2013) argue that the method of selecting $k$ should depend on the context of the clustering and frame the issue as one of obtaining an appropriate subject-matter-dependent definition of rather than a statistical estimation.

* Everitt (2011), p126
  * clusterSim
  * Continuous data?
    * Calinski and Harabasz (1974)
    * Duda and Hart (1973)
  * Steinley, D. (2006a) K-means clustering: a half-century synthesis. British Journal of Mathematical
& Statistical Psychology, 59, 1–34.  
  
* Milligan and Cooper (1984)
  + list 30

### Subs-Full


```{r SUBS_FULL, eval = F}
# K <- 20
# 
# distance <- df[, subs_f_vars[-2]] %>%
#   mutate(MEDINC = log(MEDINC),
#          n = log(n)) %>%
#   # sample_n(100) %>%
#   daisy(metric = "gower")
# 
# 
# clust_time <- system.time(
#   clusters <- run_clusters(K, distance)
# )
# 
# ch_time <- system.time(
#   chPlot <- get_CH(K, distance, clusters)
# )
# 
# # undebug(get_CH)
# 
# save(clusters, chPlot, clust_time, ch_time, K, distance, file =  "Paper Data/clusters-full-logs.rdata")
```

```{r SUBS_FULL_CH}
# load("clusters-full.rdata")
load("Paper Data/clusters-full-logs.rdata")


chPlot %>%
  gather(key = method, value = value, ch1:dif) %>%
  ggplot(aes(x = k, y = value, color = method)) +
  geom_point() +
  geom_line()



```

```{r SUBS_FULL_SSB/SST}
cls <- bind_cols(lapply(clusters, function(x) data.frame(x$cluster)))
vrat <- unlist(lapply(clusters, function(x) x$betweenss / x$totss))

data.frame(k = 1:K, var = vrat, min80 = sum(vrat < .8) + .5) %>%
  ggplot(aes(x = k, y = var)) +
  geom_point() +
  ggtitle("Between cluster variance by number of strata") +
  labs(y = "Between Cluster Variance",
       x = "Number of Strata (k)") +
  geom_line() +
  geom_vline(aes(xintercept = min80), linetype = "dashed") +
  theme_minimal() +
  scale_x_discrete(limits = c(1:K)) +
  scale_y_continuous(breaks = seq(0, 1, .1))

df$cluster_full_6 <- cls[,6]
df$cluster_full_10 <- cls[,10]
dist_full <- distance

```

### Subs-OV


```{r SUBS_OV, eval = F}
# K <- 20
# 
# distance <- df[, subs_f_vars[c(-2, -9)]] %>%
#   mutate(n = log(n)) %>% 
#   daisy(metric = "gower")
# 
# 
# clust_time <- system.time(
#   clusters <- run_clusters(K, distance)
# )
# 
# ch_time <- system.time(
#   chPlot <- get_CH(K, distance)
# )
# 
# 
# save(clusters, chPlot, clust_time, ch_time, K, distance, file =  "Paper Data/clusters-OV-logs.rdata")
```

```{r SUBS_OV_CH}
# load("clusters-OV.rdata")
load("Paper Data/clusters-OV-logs.rdata")


chPlot %>%
  gather(key = method, value = value, ch1:dif) %>%
  ggplot(aes(x = k, y = value, color = method)) +
  geom_point() +
  geom_line()



```

```{r SUBS_OV_SSB/SST}
cls <- bind_cols(lapply(clusters, function(x) data.frame(x$cluster)))
vrat <- unlist(lapply(clusters, function(x) x$betweenss / x$totss))

data.frame(k = 1:K, var = vrat, min80 = sum(vrat < .8) + .5) %>%
  ggplot(aes(x = k, y = var)) +
  geom_point() +
  ggtitle("Between cluster variance by number of strata") +
  labs(y = "Between Cluster Variance",
       x = "Number of Strata (k)") +
  geom_line() +
  geom_vline(aes(xintercept = min80), linetype = "dashed") +
  theme_minimal() +
  scale_x_discrete(limits = c(1:K)) +
  scale_y_continuous(breaks = seq(0, 1, .1))

df$cluster_OV_6 <- cls[,6]
df$cluster_OV_10 <- cls[,10]
```

```{r class_full}
df %>%
  select(cluster_full_6:cluster_OV_10) %>%
  rbind(rbind(mutate(., cluster_full_6 = 20), 
              mutate(., cluster_full_10 = 20))) %>%
  group_by(cluster_full_6, cluster_full_10) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  # mutate(p = n/sum(n)) %>%
  spread(key = cluster_full_6, value = n)

```

```{r class_OV}
df %>%
  select(cluster_full_6:cluster_OV_10) %>%
  rbind(rbind(mutate(., cluster_OV_6 = 20), 
              mutate(., cluster_OV_10 = 20))) %>%
  group_by(cluster_OV_6, cluster_OV_10) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  # mutate(p = n/sum(n)) %>%
  spread(key = cluster_OV_6, value = n)

```

```{r class_k6}
df %>%
  select(cluster_full_6:cluster_OV_10) %>%
  rbind(rbind(mutate(., cluster_full_6 = 20), 
              mutate(., cluster_OV_6 = 20))) %>%
  group_by(cluster_full_6, cluster_OV_6) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  # mutate(p = n/sum(n)) %>%
  spread(key = cluster_full_6, value = n)

```

```{r class_k10}
df %>%
  select(cluster_full_10:cluster_OV_10) %>%
  rbind(rbind(mutate(., cluster_full_10 = 20), 
              mutate(., cluster_OV_10 = 20))) %>%
  group_by(cluster_full_10, cluster_OV_10) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  # mutate(p = n/sum(n)) %>%
  spread(key = cluster_full_10, value = n)

```

\newpage

# References
```{r create_r-references}
# r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
