---
title             : "Analysis Write-up"
shorttitle        : "Analysis Write-up"

author:
  - name          : "Gleb Furman"
    affiliation   : "1"
  #   corresponding : yes    # Define only one corresponding author
  #   address       : "Postal address"
  #   email         : "my@email.com"
  # - name          : "Ernst-August Doelle"
  #   affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Who Kneads a PH.D. Bakery"
#   - id            : "2"
#     institution   : "Konstanz Business School"

# author_note: |
#   Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

#   Enter author note here.

# abstract: |
#   Enter abstract here. Each new line herein must be indented, like this line.
  
# keywords          : "keywords"
# wordcount         : "X"

# bibliography      : ["r-references.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache = F)
```

```{r load.packages, include = FALSE}
library(papaja)
library(tidyverse)

rm(list = ls())
```

```{r analysis_preferences}
# Seed for random number generation
# set.seed(42)
```

```{r data}
load("data/base data.rdata")

vars <- c("LSTATE", "LEANM", "SCHNAM", "DID", "SID", "DSID", "n", "pTotfrl",
          "Urban", "Suburban", "ToRu",
          "pELL", "pED", "pELA", "pMath", "pMin", "MEDINC", "MEDINC_SD")

# Covarites 
subs_f_vars <- c("n", "pTotfrl", "urbanicity",
                "pELL", "pED", "pELA", "pMath", "pMin", "MEDINC")

subs_ov_vars <- c("n", "pTotfrl", "urbanicity",
                "pELL", "pED", "pELA", "pMath", "pMin")

# School Level Data
df.sch <- df[,vars]


# District Level Data About students
df.dist <- list(.vars = lst("n", vars[8:17], "m"),
     .funs = lst(mean, funs(weighted.mean(., w = n)), mean)) %>%
  pmap(~ df.sch %>%
         mutate(m = n()) %>%
         group_by(DID) %>%
         summarise_at(.x, .y)) %>%
  reduce(inner_join, by = "DID")


```

```{r descriptives}
tbl_desc <- df.dist %>%
  gather(key = Variable, value = Value, n:MEDINC, m) %>%
  group_by(Variable) %>%
  summarise(M = mean(Value), SD = sd(Value))

tbl_desc <- df.sch %>%
  gather(key = Variable, value = Value, n:MEDINC) %>%
  group_by(Variable) %>%
  summarise(M = mean(Value), SD = sd(Value))  %>%
  right_join(tbl_desc, by = "Variable")



tbl_desc$Variable <- c("Number of Schools", "Median Income", "School Size", "Economically Disadvantaged", "ELA Proficiency", "English Language Learners", "Math Proficiency", "Minority Status", "Total/Free/Reduced Lunch", "Suburban", "Town or Rural", "Urban")

tbl_desc <- tbl_desc[c(1,3,2,5,7,4,6,8,9,12,10,11),]


```

```{r tbl-desc, results = "asis"}

apa_table(tbl_desc,
          caption = "Descriptives of variables",
          note = "District variables are derived as aggregate means of school variables",
          stub_indents = list(`Average Proportions` = c(4:9), Indicators = c(10:12)),
          col_spanners = list(`School` = c(2,3), `District Weighted` = c(4,5)),
          align = "lcccc",
          col.names = c("Variables","Mean", "SD", "Mean", "SD"))

```


# Cluster Analysis


## Population Frame
The population frame is composed of data from three sources: (1) the Common Core of Data (CCD), (2) publically available accountability data, and (3) the U.S. Census. The CCD is a comprehensive database housing anually collected national statistics of all public schools and districts. Accountability data was used to calculate the proportion of students within each school performing at or above proficiency in Math and ELA. Finally, local median income was obtained from the U.S. Census and was matched to each school by zipcode. Weighted means of school level variables were calculated using school size (number of students) to generate district level covariates. In this sense, district covariates describe the population of students rather than the population of schools. These are reported in Table \@ref(tab:tbl-desc)

### Covariates
Selection of covariates was driven by prior research on district and school participation behavior in RCTs (Stuart et al., 2017; Tipton et al., 2016a; Fellers, 2017). Districts and schools with higher proportions of students who are English language learners (ELL), economically disadvantaged (ED), non-White, and living in urban settings are more likely to participate, as are larger districts and schools. It is important to note, however, that some of these characteristics might also make it more likely that researchers would recruit these districts and schools in the first place. Anecdotal evidence from several research teams also suggests that schools are less willing to participate in experimental interventions for subjects in which their students are already excelling, therefore math and ELA achievement covariates were also included.

### Omitted Variable
Neighborhood median income was selected to serve as the omitted variable for several reasons. First, it is related to sevarl of the other variables and therefore likely plays a role in the likelyhood of a school participating. Second, because it comes from a non-educational data source (the census) and requires additional work to include in the population frame it is likely to be omitted in practice.

```{r dists, results = "asis", fig.cap="Comparison of covariate distributions and their log transformations."}
to_z <- function(x) (x - mean(x))/sd(x)

dist_plots <- df[, c("DSID",subs_f_vars[-2])] %>%
  mutate(n_log = log(n),
         MEDINC_log = log(MEDINC/1000),
         MEDINC = MEDINC / 1000)

dist_plots %>%
  gather(key = var, value = value, n, n_log, MEDINC, MEDINC_log) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~ var, scales = "free") +
  theme_apa()

# dist_plots %>%
#   gather(key = var, value = value, n:MEDINC_log, -urbanicity) %>%
#   ggplot(aes(x = value, color = urbanicity)) +
#   geom_density() +
#   facet_wrap(~ var, scales = c("free"))
# 
# dist_plots %>%
#   ggplot(aes(x = pELA, y = pMath, color = n_log)) +
#   geom_point(alpha = .5) +
#   facet_wrap(~urbanicity)
# 
# dist_plots %>%
#   ggplot(aes(x = pELL, y = pMin, color = n_log)) +
#   geom_point(alpha = .5) +
#   facet_wrap(~urbanicity)
# 
# dist_plots %>%
#   ggplot(aes(x = MEDINC_log, y = pED, color = n_log)) +
#   geom_point(alpha = .5) +
#   facet_wrap(~urbanicity)
```

```{r dists2, fig.cap="Distributions of the remaining continuous covariates."}
dist_plots %>%
  gather(key = var, value = value, pED:pMin) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~ var, scales = "free") +
  theme_apa()

```

### Variable Transformations
Log-transformation was used on school size (number of students) and median income. This is done to allow proportional comparisons at the extremes of the distributions (Hennig and Liao - 2013). For instance, the difference between two schools with 4000 and 3000 students should be weighed as much as the difference between two schools with 400 and 300 students when generating clusters. Figure \@ref(fig:dists) displays a comparison of the distribution of these variables and hteir logs. Figure \@ref(fig:dists2) displays the distribution of the remaining variables.


## SUBS

Stratification using balanced sampling (SUBS) was performed prior to simulation because the group of schools in each strata would be static across conditions except where the balancing model is manipulated. The set of covariates in both the full model (SUBS-F) and the omitted variable model (SUBS-OV) include binary indicator variables

### Number of Clusters
Selecting the number of clusters, $k$, is one of the most difficult problems in cluster analysis (Steinley, 2006). To date, the most extensive investigation of methods for determining $k$ was conducted by Milligan and Cooper (1985) who analyzed 30 methods. However, aside from the limited generalizability of this study, many methods are also innapropriate in the context of non-hierarchical and thus do not support k-means clustering. Hennig and Liao (2013) argue that the method of selecting $k$ should depend on the context of the clustering and frame the issue as one of obtaining an appropriate subject-matter-dependent definition of rather than a statistical estimation.

* Everitt (2011), p126
  * clusterSim
  * Continuous data?
    * Calinski and Harabasz (1974)
    * Duda and Hart (1973)
  * Steinley, D. (2006a) K-means clustering: a half-century synthesis. British Journal of Mathematical
& Statistical Psychology, 59, 1–34.  
  
* Milligan and Cooper (1984)
  + list 30

### Subs-Full


\newpage

# References
```{r create_r-references}
# r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
