---
title             : "Analysis Write-up"
shorttitle        : "Analysis Write-up"

author:
  - name          : "Gleb Furman"
    affiliation   : "1"
  #   corresponding : yes    # Define only one corresponding author
  #   address       : "Postal address"
  #   email         : "my@email.com"
  # - name          : "Ernst-August Doelle"
  #   affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Who Kneads a PH.D. Bakery"
#   - id            : "2"
#     institution   : "Konstanz Business School"

# author_note: |
#   Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

#   Enter author note here.

# abstract: |
#   Enter abstract here. Each new line herein must be indented, like this line.
  
# keywords          : "keywords"
# wordcount         : "X"

# bibliography      : ["r-references.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache = F)
```

```{r load.packages, include = FALSE}
library(papaja)
library(tidyverse)

rm(list = ls())
```

```{r analysis_preferences}
# Seed for random number generation
# set.seed(42)
```

```{r data}
load("data/base data.rdata")

vars <- c("LSTATE", "LEANM", "SCHNAM", "DID", "SID", "DSID", "n", "pTotfrl",
          "Urban", "Suburban", "ToRu",
          "pELL", "pED", "pELA", "pMath", "pMin", "MEDINC", "MEDINC_SD")

# Covarites 
subs_f_vars <- c("n", "pTotfrl", "urbanicity",
                "pELL", "pED", "pELA", "pMath", "pMin", "MEDINC")

subs_ov_vars <- c("n", "pTotfrl", "urbanicity",
                "pELL", "pED", "pELA", "pMath", "pMin")

# School Level Data
df.sch <- df[,vars]


# District Level Data
df.dist <- select(df.sch, -SCHNAM, -LSTATE, -LEANM, -SID, -MEDINC_SD) %>%
  group_by(DID) %>%
  mutate(m = n()) %>%
  summarise_if(is.numeric, mean) %>%
  ungroup()


df.dist.w <- list(.vars = lst("n", vars[8:17], "m"),
     .funs = lst(mean, funs(weighted.mean(., w = n)), mean)) %>%
  pmap(~ df.sch %>%
         mutate(m = n()) %>%
         group_by(DID) %>%
         summarise_at(.x, .y)) %>%
  reduce(inner_join, by = "DID")


df.dist$dif <- df.dist.w$dif <- apply((df.dist[,-1] - df.dist.w[,-1])/apply(df.dist[,-1], 2, function(x) max(x) - min(x)), 1, function(x) mean(abs(x)))



```

```{r descriptives}
tbl_desc <- df.dist %>%
  gather(key = Variable, value = Value, n:MEDINC, m) %>%
  group_by(Variable) %>%
  summarise(M = mean(Value), SD = sd(Value))

tbl_desc <- df.dist.w %>%
  gather(key = Variable, value = Value, n:MEDINC, m) %>%
  group_by(Variable) %>%
  summarise(M = mean(Value), SD = sd(Value)) %>%
  right_join(tbl_desc, by = "Variable")

tbl_desc <- df.sch %>%
  gather(key = Variable, value = Value, n:MEDINC) %>%
  group_by(Variable) %>%
  summarise(M = mean(Value), SD = sd(Value))  %>%
  right_join(tbl_desc, by = "Variable")



tbl_desc$Variable <- c("Number of Schools", "Median Income", "School Size", "Economically Disadvantaged", "ELA Proficiency", "English Language Learners", "Math Proficiency", "Minority Status", "Total/Free/Reduced Lunch", "Suburban", "Town or Rural", "Urban")

tbl_desc <- tbl_desc[c(1,3,2,5,7,4,6,8,9,12,10,11),]


```

```{r tbl_desc, results = "asis"}

apa_table(tbl_desc,
          caption = "Descriptives of variables",
          note = "District variables are derived as aggregate means of school variables",
          stub_indents = list(`Average Proportions` = c(4:9), Indicators = c(10:12)),
          col_spanners = list(`School` = c(2,3), `District Weighted` = c(4,5), `District Unweighted` = c(6,7)),
          align = "lcccccc",
          col.names = c("Variables","Mean", "SD", "Mean", "SD", "Mean", "SD"))

```


# Cluster Analysis


## Population Frame
The population frame is composed of data from three sources: (1) the Common Core of Data (CCD), (2) publically available accountability data, and (3) the U.S. Census. The CCD is a comprehensive database housing anually collected national statistics of all public schools and districts. Accountability data was used to calculate the proportion of students within each school performing at or above proficiency in Math and ELA. Finally, local median income was obtained from the U.S. Census and was  matched to each school by zipcode. School level data was aggregated to get district level variables. These are reported in Table \@ref(tab:tbl_desc)



### SUBS
Stratification using balanced sampling (SUBS) was performed prior to simulation because the group of schools in each strata would be static across condtions except where the balancing model is manipulated. The set of covariates in both the full model (SUBS-F) and the omitted variable model (SUBS-OV) include binary indicator variables ()


```{r SUBS_FULL}
library(parallel)
library(cluster)

K <- 20

run_clusters <- function(K, distance) {

  # Calculate the number of cores
  no_cores <- min(K, detectCores() - 1)

  # Initiate cluster
  cl <- makeCluster(no_cores)

  clusters <- parLapply(cl, 1:K, function(x, y) kmeans(x = y, centers = x), distance)

  stopCluster(cl)

  # clusters <- bind_cols(clusters)
  #
  # names(classign) <- paste("k", 1:K, sep = "")

  return(clusters)
}


distance <- daisy(x = df[, subs_f_vars[-2]], metric = "gower")

clusters <- run_clusters(K, distance)

save(clusters, file =  "Paper Data/clusters-omitted-totfrl.rdata")

load("Paper Data/clusters-full.rdata")

cls <- bind_cols(lapply(clusters, function(x) data.frame(x$cluster)))
vrat <- unlist(lapply(clusters, function(x) x$betweenss / x$totss))



data.frame(k = 1:K, var = vrat, min80 = sum(vrat < .8) + .5) %>%
  ggplot(aes(x = k, y = var)) +
  geom_point() +
  ggtitle("Between cluster variance by number of strata") +
  labs(y = "Between Cluster Variance",
       x = "Number of Strata (k)") +
  geom_line() +
  geom_vline(aes(xintercept = min80), linetype = "dashed") +
  theme_minimal() +
  scale_x_discrete(limits = c(1:K)) +
  scale_y_continuous(breaks = seq(0, 1, .1))

# ggsave("Params/Scree1.png", dpi = 500, width = 5, height = 5)


# df$cluster_full <- df_cluster$k6


```

\newpage

# References
```{r create_r-references}
# r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
